### 2015-01-04 2014/nature-review/deepreview.tex

\documentclass[]{article}
\usepackage{epsfig}
\usepackage{graphicx}
\parskip 0.1in
\parindent 0.0in
\setcounter{secnumdepth}{1}
\title{Deep Learning}

\author{Yann LeCun, Yoshua Bengio, Geoffrey Hinton}

\begin{document}

\maketitle

\section{Preface}

Deep neural networks that learn many layers of feature detectors have dramatically improved the state-of-the-art for
speech recognition, object recognition, object detection, predicting the activity of drug molecules, and many other
tasks. Deep learning discovers intricate structure in large datasets by using the backpropagation algorithm to compute
error-derivatives for the weights that define each feature detector and then repeatedly changed these weights in
proportion to their derivatives averaged over small batches of training examples.  Deep recurrent neural networks, in which
each layer also receives input from its own state at the previous time-step, can now generate good captions for images
and are competitive for machine translation.

\section{Introduction}

Deep learning is currently making major advances in solving problems that have resisted the best attempts of the
Artificial Intelligence community for many years. Deep learning is very good at discovering intricate structure in
large, high-dimensional datasets and is therefore widely applicable in all of the sciences. For example, it beats all
previous machine learning techniques at predicting the activity of potential drug molecules~\cite{Dahl}, detecting new
fundamental particles~\cite{??}, classifying galaxies~\cite{??}, or predicting the effects of mutations in non-coding
DNA on gene expression and disease~\cite{Frey}.

So far, most applications of deep learning use feed-forward neural networks, but for tasks that involve sequential
inputs it is often better to use recurrent neural networks. These are given one term from an input sequence at each time
step and their recurrent internal ``hidden'' state depends on the whole history of the input sequence. Recurrent
networks are very good at predicting the next character in a string\cite{Sutskever} or the next word in a
sequence\cite{Mikolov} but they can also be used for more complex tasks.  After reading an English sentence one word at
a time, for example, an English``encoder'' network can be trained so that the final state vector of its hidden neurons
is a good representation of the thought expressed by the sentence.  This thought vector can then be used as the initial
hidden state of a trained French ``decoder'' network which outputs a probability distribution for the first word of the
French translation. If a particuar first word is chosen and provided as input to the decoder network it will then output
a probability distribution for the second word of the translation and so on until a full stop is chosen~\cite{Vinyals, Montreal}.

This rather naive way of performing machine translation has quickly become competitive with the state-of-the-art and
this raises serious doubts about whether understanding a sentence requires anything like internal symbolic expressions
that are manipulated by binding their symbols to the variables in judiciously chosen rules of inference.  The
encoder-decoder translation networks do not work by using internal rules of inference in which variables get bound to
symbols. They just use big state vectors, big weight matrices and scalar non-linearities to get the job done.  This
suggests that deriving conclusions using only the form of the premises is a powerful technique that people can
eventually master but it is not a good model of the way we normally reason or understand language. Recurrent neural
networks that represent thoughts as big vectors are more compatible with the view that everyday reasoning involves many
simultaneous analogies that each contribute plausibility to a conclusion\cite{metaphors}.

This review is organized chronlogically starting with the introduction of the backpropagation algortihm in the
1980's. After describing some of its early successes, we explain the actual and perceived limitations that prevented it
from living up to its early promise for so many years and how these limitatons were eventually overcome.

\section{Pattern recognition and the back-propagation algorithm}

Patterns can be classified by first converting each raw input vector into a vector of feature activations and then
collecting evidence for each possible class by computing a weighted sum of the features. If good features can be
designed by hand, it is relatively easy to learn class-specific weights that optimize
discrimination~\cite{perceptrons}. Widely-used example of such linear classification methods are logistic
 regression and linear support vector machines~\cite{logisticregression,svm}.

However, even with good hand-crafted features, many interesting tasks cannot be solved by linear classifiers. A common
method to get around this problem is to transform the feature vector into a higher dimensional vector through a
non-linear mapping, exploiting the fact that collections of vectors are more likely to be easily spearated when they
are spread around a high-dimensional space~\cite{cover}. One clever way to do so is to use a ``kernel function'' that
measures the similarity between two vectors.  Using the kernel function, any input vector, ${\bf x}_i$ in the training
set can be used to define a feature whose activation for a new input vector, ${\bf x}_{\rm new}$ is just the
similarity between ${\bf x}_i$ and ${\bf x}_{\rm new}$ . Given a kernel function, it is possible to jointly optimize
the choice of training inputs used to define features and the class-specific weights associated with the
features\cite{}. The training inputs used to define features are called ``support vectors'' and Support Vector
Machines\cite{Vapnik} have been very successful, especially for tasks with a relatively small number of training
examples in which features are individually informative. They do not, however, perform well for tasks such as object
recognition in which the input space is high-dimensional and the input-output mapping is so
complex that computing similarities with a simple kernel function is ineffective.

Problems such as image and speech recognition require the input-output function to be insensitive to irrelevant
variations of the input, such as variations in position, orientation or illumination of an object, or variation in
pitch or accent of speech, while being very sensitive to particular minute variations, such as the difference between
a white wolf and a samoyed dog. These tasks are better handled with multiple layers of progressively more
complicated feature detectors each of which is defined in terms of the simpler feature detectors in the previous
layers.

From the earliest days of pattern recognition\cite{selfridge, rosenblatt} researchers aimed to replace hand-engineered
features by multiple layers of learned features.  This was eventually achieved by using the backpropagation algorithm
which was discovered independently by several different groups in the 1970s and 1980s\cite{Werbos, Parker, Lecun,
  Rumelhart, schmidreference}.  Backpropagation is simply an application of chain rule, which yields an efficient way
of computing the gradient of a smooth objective function with respect to all of the weights in a layered network. The
objective function typically measures the discrepancy between the actual output of the network on a given training
case and the target output for that case.  The key insight is that the error derivatives with respect to the {\it
  outputs} of a layer can be computed by working backwards from the error derivatives with respect to the outputs of
the subsequent layer. Once this backwards recursion has been used to compute error derivatives for the outputs of all
layers, it is easy to compute error derivatives for the parameters (or weights) of each layer (see box 1).

%%%%%%% GEOFF %%%%%%%%%%
GEOFF:  Do box explaining backprop equatons
%%%%%%%%%%%%%%%%%%%%%%

\section{Stochastic gradient descent}

Backpropagation provides an efficient way of computing gradients for all of the weights in a layered network on a
single training case, but there are still many ways to make use of these gradients. The most obvious method is to
average the gradients over all of the cases in the training set and then to change each weight in proportion to its
negative gradient. When the training set contains millions of cases and is highly redundant, however, this is
needlessly slow. It is much faster to compute the average gradient on a small ``mini-batch'' of randomly chosen
training cases and to update each weight by this stochastic estimate of its full gradient times a learning rate which
decays as learning proceeds. Stochastic gradient descent can be improved by using the history of the individual weight
gradients to adapt the learning rate for each weight\cite{bottou-bousquet} or to provide additional increments that
depend on earlier gradients\cite{} but these techniques are still the subject of controversy and are beyond the scope
of this review.  However, it is now widely accepted that some form of stochastic gradient descent is the best bet for
large, redundant datasets.

\section{Distributed representations}

The hidden layers of a multi-layer neural network learn to represent the network's inputs in a way that
makes it easy to predict the target outputs. This is nicely demonstrated by training a multi-layer neural network to
predict the next word in a sequence from a local context of earlier words. Each context word is presented to the network
as a one-of-N vector in which one component has a value of $1$ and the rest are $0$. The output layer has one unit per
word and is forced to produce a probability distribution over words by using the ``softmax'' function:
\begin{equation}
p_j = \frac{exp(z_j)}{sum_i exp(z_i)}
\end{equation}
where $z_j$ is the total input received by output unit $j$. The hidden layers use either the logistic or tanh non-linearity:
\begin{equation}
y_j = \frac{1}{1+ exp(z_j)}, \ \ \ y_j = \frac{}{}
\end{equation}

The connectivity of the network is restricted so that each of the one-of-N input vectors must first be converted to a
``word-vector'' before being used by subsequent layers to give a high probability to the target next word.  The network
learns word-vectors that contain many active components each of which can be interpreted as a separate feature of the
word. This was first demonstrated\cite{RHW} using sequences of three words, such as (Colin, Mother, Victoria), derived
from two family trees. The learned word vector for Colin contained ``semantic'' features representing which family tree
Colin was in, what branch of that tree he was in, and what generation he was\cite{hinton86}. These semantic features
were not explicitly present in the input.  They were discovered by the learning procedure as a good way of factorizing
the structured relationships between the input and output symbols into multiple ``micro-rules''.  For example, given
Colin's generation and the learned feature of Mother that the output should be one generation higher than the input, we
can already restrict the answer to people in a specific generation. The ability of deep learning to discover many such
micro-rules allows it to generalize well to novel combinations of individually familiar context words.

The family tree demonstration used a small, clean dataset in which the regularities could be captured by micro-rules
that were never violated and the features were easy to interpret, but learned word-vectors also work very well when the word
sequences come from a large corpus of real text and the individual micro-rules are unreliable\cite{NPLM} . When trained
to predict the next word in a news story, for example, the learned word vectors for Monday and Tuesday are very similar,
as are the word vectors for Vietnam and Iraq.  Vector representations of words learned from text are now very widely
used in natural language applications\cite{Collobert, Mikolov, Socher?}.

The issue of representation lies at the heart of the debate between the logic-inspired and the neurally-inspired
paradigms for cognition. In the logic-inspired paradigm, an instance of a symbol is something whose only property is
that it is either identical or non-identical to other symbol instances. It has no internal structure that is
relevant to its use. By contrast, neural networks use large activity vectors instead of symbols and parallel matrix
operations followed by non-linearities instead of formal rules of inference. With sufficient data, computational power
and depth, the combination of backpropagation for computing gradients and stochastic gradient descent for updating the
weights on connections has proved to be very good at learning appropriate distributed representations. These
representations allow a lot of parallel computation at the feature level so that complex tasks can be solved in a very
small number of sequential steps.

\section{Convolutional neural networks}

%%%%%%%% YANN %%%%%%%

YANN: can you summarize convnets and their use for reading cheques in about 350 words?

%%%%%%%%%%%%%%%%%%%%

\section{Recurrent Neural Networks}

When backpropagation was first introduced, its most exciting use was for training recurrent neural networks. These
receive input vectors at each time-step and have one or more layers of hidden units that, in addition to having
connections to higher layers or ouput units, also have connections that allow the neurons within a layer to receive
weighted input from those same neurons at the previous time-step.  The gradient for a recurrent weight, $w_{ij}$ that
allows neuron $i$ at time $t$ to influence the activity of neuron $j$ at time $t+1$ can be computed by
backpropagation-through-time (BPTT) of the error derivative for $j$ at time $t+1$. The full gradient for $w_{ij}$ on a whole
input sequence is just the sum of its gradients at each time-step.

Recurrent neural networks are very powerful dynamical systems, but training them with BPTT proved to be problematic,
because the backpropagated gradients either grow or shrink at each time step, so over many time-steps they typically explode
or vanish\cite{Bengio, schmdhuber}.

\section{Overcoming the limitations of backpropagation in deep networks}

In the 1990s, the machine learning community decided that backpropagation in deep neural networks did not work well in
practice and it was largely abandoned.  At the time, people believed that learning many layers of non-linear feature
detectors with no prior knowledge was much too difficult for a simple gradient descent procedure and that the networks
were getting trapped in poor local minima. We now know that the main reasons for the limited success of backpropagation
in deep networks were quite different: The labeled datasets were far too small, the computers were far too slow, and the
weights were not initialised sensibly.

The obvious weakness of gradient descent is that it can get trapped in poor local minima, but in high-dimensional
weight-spaces this is much less common than getting stuck close to saddle points where the error surface curves upwards
in most directions but downwards in a few. Recent empirical\cite{Bengio} and theoretical results\cite{Yann} strongly
suggest that in deep networks almost all of these saddle points give very similar performance so it does not matter
which one we find.  This goes a long way towards explaining why stochastic gradient descent works so well for deep
neural networks.

Interest in deep feed-forward networks was revived in 2005\cite{IJCAI} by the introduction of an unsupervised learning
procedure that could create layers of feature detectors without requiring labeled data. The objective in learning each
layer of feature detectors was to be able to reconstruct or model the activities of feature detectors (or raw inputs) in the
layer below.  By ``pre-training''  several layers of progressively more complex feature detectors using this reconstruction
objective, the weights of a deep network could be initialized to sensible values.  A final layer of output units could
then be added to the top of the network and the whole deep system could be fine-tuned using standard
backpropagation\cite{Science}. This worked remarkably well for recognizing handwritten digits especially when the amount
of labeled data was very limited\cite{BengioNips06}.

The first major application of this pre-training approach was in speech recognition and it was made possible by the
advent of fast Graphics Processing Units (GPUs) that were convenient to program.  In 2009, the approach was used to map
short temporal windows of coefficients extracted from a soundwave to a set of probabilities for the various fragments of
speech that might be represented by the frame in the center of the window.  It achieved record-breaking results on a
standard speech recognition benchmark that used a small vocabulary\cite{TIMITpaper} and was quickly developed to give
record-breaking results on a large vocabulary task\cite{Dahl}.  By 2012, versions of the deep net from 2009 were being
developed by many of the major speech groups\cite{SPS} and were already being deployed in Android phones.  For smaller
datasets, unsupervised pre-training leads to significantly better generalization, but once deep
learning had been rehabilitated, it turned out that large labeled datasets did not need the pre-training stage, provided the
scales of the initial random weights were carefully chosen to avoid very large or very small initial gradients.

Interest in recurrent networks (which are deep in time) was revived by ``Long Short Term Memory'' networks (LSTMs) that
use special hidden units whose natural behaviour is to remember inputs for a long time\cite{Hochreiter}.  A
linear ``memory cell'' interacts with other units via gated connections whose weights are multiplied by the state of a
logistic gating unit that has an activation between $0$ and $1$. The input vector has little effect on the state of a
memory cell if the cell's input gate has a low activation so memory cells can learn to selectively ignore the input.  Similarly,
a memory cell has little effect on the output if its output gate has low activation. Each memory cell also has a
connection to itself at the next time-step that has a weight of 1 so it tends to copy its own real-valued state, but
this connection is gated by a ``keep gate'' so the stored value is quickly eroded unless the keep-gate has an activation
near $1$.  The gates themselves have their activities determined by weighted inputs that they receive from the input
vectors and all of the memory cells.

When LSTMs were first described in English\cite{HochSchmid} their complexity was off-putting and, despite their
impressive ability to learn very long-range temporal structure on toy problems, they were largely ignored until they
achieved excellent results in a competition to read cursive hand-writing\cite{Graves}. They have subsequently proved to
be a very effective for tasks that have sequential input.  They are now better than deep feed-forward networks for
speech recognition\cite{gravestimit} and unlike feed-forward nets they can implement an entire speech recognition system
that goes all the way from coefficents derived from the soundwave to the sequence of characters in the transcription~\cite{graves}.
They are also what is currently used for the encoder and decoder networks that perform so well at machine translation.

It is not yet clear which aspects of the complicated architecture of LSTMs are really necessary. Memory cells are a good
way of allowing the backpropagated error derivatives to neither explode nor vanish, but they are not the only way.  For
example, recurrent nets with rectified linear hidden units that use the non-linear function $y=max(0,x)$ are much
simpler than LSTMs and if the recurrent weight matrix is initialized to be close to the identity matrix they work about
as well as LSTMs both for toy problems involving very long-range temporal structure and for real tasks like predicting
the next word in text\cite{LeHinton}.

\section{Object recognition and detection with deep convolutional networks}

When deep convolutional networks were applied to a dataset of about a million images from the web that contained a
thousand different classes they achieved spectacular results, almost halving the error rates of the best computer vision
sytems~\cite{kriz}.  Some of the images and the network's predictions are shown in figure ??.  The major factors in this
success were the large labeled dataset~\cite{FeiFe}, the use of fast GPUs, and the convolutional early layers. were.
Rectified linear units made the training much faster and the fully connected final layers were prevented from
overfitting the peculiarities of the training set by randomly omitting hidden units during
training\cite{dropoutJMLR}. Deep convolutional nets have quickly replaced the carefully engineered pipelines full of
hand-designed features developed by the computer vision community and they are now deployed in many computer vision
applications, including image tagging, object detection, face recognition, facial expression recognition and scene parsing.

%%%%%%%% YANN %%%%%%%
YANN: Add about 300 words plus another part of the figure showing detected parts of people. 
%%%%%%%%%%%%%%%%%%%%

\section{Language processing with deep networks}

%%%%%%%% YOSHUA %%%%%%%
YOSHUA: Add about 300 words plus t-sne figure of words and/or phrases.

The intro already explained the simple version of translation using encoder and decoder nets and I  suggest we  save
your ``translating with attention'' for introducing attention in the 
the future section.

After your stuff on language, I would like to include some version of the paragraph below along with a figure.
%%%%%%%%%%%%%%%%%%%%%%%

When translating between languages, the hidden units of the decoder network that generates possible translations are
initalized with a ``thought vector'' which is the final hidden state of an encoder network.  But it is also possible to
initialize a decoder network using a linear transformation of the last hidden layer of a deep convolutonal network that
has already been trained to recognize objects in images. The decoder network can then be trained by BPTT on a dataset of
captioned images~\cite{NIC,StanfordNIC,more?} and it produces remarkably good captions (see figure ??).

\section{Deep reinforcement learning}

Learning which action will maximize the total future reward given a state of the world is difficult when the rewards are
delayed and the world has a very large number of possible states. One impressive recent development is a single system
that can learn to play many different video games~\cite{NatureAtari}. The system uses a deep convolutional network to
map from the pixels on the screen at several previous time steps to the expected future reward of each possible action
and its only inputs are the pixels and the current score in the game. The learning works by minimizing the difference
between the system's current estimate of future reward and a noisy but less biased estimate that is obtained by using
the same deep network to estimate the future reward of the next state, where the next state is chosen stochastically
based on the predicted future rewards of all the currently legal actions.

\section{Learning to process symbols}

Deep learning is now the method of choice for many perceptual tasks and is making major inroads in natural language
processing where a huge amount of messy commonsense knowledge is required. Rapid progress is also being made in applying deep

learning, in particular BPTT, to relatively pure symbol processing tasks.  An impressive recent example is the ``Neural
Turing Machine'' which is a recurrent neural network that can choose when and what to write to or read from an external
memory~\cite{NTM}. Among other things, this can learn to ouput a sorted list of symbols when its input consists of an
unsorted sequence in which each symbol is accompanied by a real value that indicates its priority in the list.  To
achieve this it must learn a sorting program purely from examples of the input and output sequences.  A different system
that also uses BPTT can learn to map the sequence of characters in a simple program written in Python to the seqeunce of
characters that the program would print~\cite{Woj}. Rather surprisingly, BPTT can learn about {\it for loops} and can
also learn to perform addition and multiplication of strings contining several decimal digits. Achieving this kind of
behaviour seemed ludicrously optimistic when BPTT was first proposed.

\section{The future of deep learning}

Human vision is an active process in which intelligent fixations are used to ensure that only a tiny fraction of the optic array is ever processed at high resolution. We expect much of the future progress in vision, especially for video understanding or robotics, to come from systemsthat are trained end-to-end and combine deep convolutional networks with recurrent nets that use reinforcement learning to decide where to look\cite{}.  Such systems already outperform passive vision systems\cite{ba\&mnih}.  Similarly, when understanding sentences or whole documents we expect future systems to learn strategies for selectively attending to one part at a time\cite{Yoshua translation}. We also expect that in the next few years deep recurrent neural networks will become far better at understanding the meaning of documents, including the way in which the thoughts expressed by the sentences are related to one another.

This review has focused on supervised learning because of its recent achievments and we have ignored much
interesting work on unsupervised learning procedures for deep neural networks~\cite{deep-boltz, wake-sleep, google-cat, denoising-auto, ranzato-pami, more recent}. In the long run, however, we believe unsupervised learning will be essential for allowing deep networks to generalize well by creating internal representations that untangle the many different factors that interact to produce images, speech  and documents~\cite{Yann, Yoshua: untangling papers}.

Finally, we believe that the recent successes of deep learning using stochastic gradient descent are only the beginning. 
End-to-end learning systems that have many parameters and few built-in assumptions can make use of more data and more compute power far more easily than systems that require hand-engineering of domain-specific knowledge.

\section{Key references}
{\it Quite a few of these are very recent and have not yet been
  published through a peer-review so we are giving their arxiv
  references}.

\subsection{Back-propagation and Stochastic Gradient Descent}
Rumelhart, D. E., Hinton, G. E., \& Williams, R. J. (1986)\\ Learning
representations by back-propagating errors.\\ {\it Nature}, {\bf 323}, 533-536.

LeCun, Y., Boser, B., Denker, J. S., Henderson, D., Howard, R.E., Hubbard, W., \& Jackel, L.D.\\ 
Backpropagation Applied to Handwritten Zip Code Recognition\\
Neural Computation, 1(4):541-551, Winter 1989

\subsection{Distributed representations}
Hinton, G. E., McClelland, J. L., \& Rumelhart, D. E. (1986)\\ Distributed
representations.\\ In Rumelhart, D. E. and McClelland, J. L., editors, {\it
  Parallel Distributed Processing: Explorations in the Microstructure of
  Cognition. Volume 1: Foundations}, MIT Press, Cambridge, MA. pp 77-109.

Bengio, Y., Ducharme, R., \& Vincent, P. (2001)\\ A neural probabilistic
language model.\\ In {\it NIPS 2000}, pages 932--938.

%% Collobert, R., Weston, J., Bottou, L., Karlen, M., Kavukcuoglu, K., Kuksa, P. (2011)\\
%% Natural language processing (almost) from scratch.\\
%% {\it The Journal of Machine Learning Research} 12, pages 2493-2537

Mikolov, T., Sutskever, I., Chen, K., Corrado, G.S., \& Dean J.\\
Distributed representations of words and phrases and their compositionality.\\
In {\it NIPS 2013}, pages 3111-3119.

Montufar, G.~F., Pascanu, R., Cho, K., and Bengio, Y. (2014).  On the number of
linear regions of deep neural networks.\\ In {\it NIPS'2014}.

\subsection{The apparent limitations of backpropagation in deep networks}

Bengio, Y., Simard, P., \& Frasconi, P. (1994)\\ Learning long-term dependencies
with gradient descent is difficult.\\ {\it IEEE Transactions on Neural
  Networks}, 5(2), 157-166.

%% Dauphin, Y., Pascanu, R., Gulcehre, C., Cho, K., Ganguli, S., and Bengio,
%% Y.\\ 
%% Identifying and attacking the saddle point problem in high-dimensional
%% non-convex optimization.\\ In {\it NIPS'2014}.

\subsection{Deep unsupervised learning}
Hinton, G., Osindero, S., \& Teh, Y. W. (2006)\\ A fast learning algorithm for
deep belief nets.\\ {\it Neural Computation}, 18(7), 1527-1554.

Erhan, D., Bengio, Y., Courville, A., Manzagol, P. A., Vincent, P., \& Bengio,
S. (2010)\\ Why does unsupervised pre-training help deep learning?\\ {\it The
  Journal of Machine Learning Research}, {\bf 11}, 625-660.

%% Mesnil, G., Dauphin, Y., Glorot, X., Rifai, S., Bengio, Y., Goodfellow, I.,
%% Lavoie, E., Muller, X., Desjardins, G., Warde-Farley, D., Vincent, P.,
%% Courville, A., and Bergstra, J. (2011).\\ Unsupervised and transfer learning
%% challenge: a deep learning approach.\\ In {\em JMLR W\&CP: Proc. Unsupervised
%%  and Transfer Learning\/}, volume~7.

\subsection{Deep speech recognition}

%% Mohamed, A. R., Dahl, G. E., \& Hinton, G. (2012)\\ Acoustic modeling using deep
%% belief networks.\\ {\it IEEE Transactions on Audio, Speech, and Language
%%  Processing}, 20(1), 14-22.

Hinton, G., Deng, L., Yu, D., Dahl, G. E., Mohamed, A. R., Jaitly, N. \&
Kingsbury, B. (2012)\\ Deep neural networks for acoustic modeling in speech
recognition: The shared views of four research groups.\\ {\it Signal Processing
  Magazine}, IEEE, 29(6), 82-97.

\subsection{Deep convolutional networks for image understanding}
LeCun, Y., Bottou, L., Bengio, Y., \& Haffner, P. (1998)\\ Gradient-based
learning applied to document recognition.\\ {\it Proceedings of the IEEE},
86(11), 2278-2324.

%% Jain, V., Seung, H.S., \& Turaga S.C. (2010)\\
%% Machines that learn to segment images: a crucial technology for connectomics\\
%% {\it Current opinion in neurobiology} 20 (5), 653-666.

Krizhevsky, A., Sutskever, I., \& Hinton, G. E. (2012)\\ Imagenet classification
with deep convolutional neural networks.\\ In Advances in Neural Information
Processing Systems (1097-1105).

%% Farabet, C., Couprie, C., Najman, L. \& LeCun, Y. (2013)\\
%% Learning Hierarchical Features for Scene Labeling,\\
%% IEEE Transactions on Pattern Analysis and Machine Intelligence, August 2013.

%% Sermanet, P., Eigen, D., Zhang, X., Mathieu, M., Fergus, R., \& LeCun,
%% Y. (2013)\\ Overfeat: Integrated recognition, localization and detection using
%% convolutional networks.\\ {\it Proceedings of ICLR 2014}, arXiv:1312.6229.

Tompson, J., Goroshin, R., Jain, A., LeCun, Y., \& Bregler, C. (2014)
Efficient Object Localization Using Convolutional Networks. arXiv:1411.4280.

\subsection{Deep recurrent neural networks for text understanding and machine translation}

Graves, A., Liwicki, M., Fernández, S., Bertolami, R., Bunke, H., \&
Schmidhuber, J. (2009)\\ A novel connectionist system for unconstrained
handwriting recognition.\\ {\it IEEE Transactions on Pattern Analysis and
  Machine Intelligence}, 31(5), 855-868.

%% Devlin, J., Zbib, R., Huang, Z., Lamar, T., Schwartz, R., \& Makhoul,
% J. (2014)\\ Fast and robust neural network joint models for statistical machine
%% translation.\\ In {\it 52nd Annual Meeting of the Association for Computational
%%   Linguistics}, Baltimore, MD, USA.

Sutskever, I., Vinyals, O., \& Le, Q. V. (2014)\\ Sequence to sequence learning
with neural networks.\\ {\it arXiv preprint}, arXiv:1409.3215.

Bahdanau, D., Cho, K., \& Bengio, Y. (2014)\\ Neural machine translation by
jointly learning to align and translate.  {\it arXiv preprint}, arXiv:1409.0473.

Vinyals, O., Toshev, A., Bengio, S., \& Erhan, D. (2014)\\ Show and Tell: A
Neural Image Caption Generator.\\ {\it arXiv preprint}, arXiv:1411.4555.

%% Weston, J., Chopra, S., Bordes, A. (2014)\\
%% Memory Networks. arXiv:1410.3916.

\subsection{Deep reinforcement learning}
Mnih, V., Kavukcuoglu, K., Silver, D., Graves, A., Antonoglou, I., Wierstra, D.,
\& Riedmiller, M. (2013)\\ Playing Atari with deep reinforcement
learning.\\ {\it arXiv preprint} arXiv:1312.5602.

\newpage

\section{Boxes}

A box for the forward propagation and backpropagation equations.

\section{Figures}

1. Diagram showing:\\ a) A feedforward neural net.\\ b) A recurrent neural net
expanded in time.\\ c) A convolutional neural net.

2. Maps showing:\\ a) that word vectors learned by backpropagation are very good
at capturing the meanings of the words.\\ b) that the thought vectors extracted
by a translation model are very good at capturing the thoughts expressed by
sentences.

3. A diagram showing how unsupervised learning can be used to pre-train a deep
feeedforward neural net.

4. Some images with:\\ a) The opinions of a deep convolutional net about the
classes of the prominent objects.\\ b) Several alternative captions generated by
using the output of the convolutional net to determine the initial hidden state
of a decoder recurrent net.

5. Some images showing the localizations of parts of a person found by a deep
convolutional neural net.\\

\end{document}